% !TEX root = ../Dissertation.tex
% !TEX spellcheck = en_US

% ABSTRACT goes here -- Maximum 350 words!

	Modern and emerging telecommunication networks employ optical technology in the backbone. Providing  high-speed connectivity, high bandwidth and reliable networks, connection failures within backbones may lead to social and economic damages which may be catastrophic. Therefore, survivability is a critical aspect for the success and the delivery of data in achieving end to end connectivity requirements in optical networks. The primary focus of this research is to investigate the effectiveness of two traditional survivable link disjoint routing algorithms, Bhandari's algorithm and iterative Djikstra's algorithm in order to compare them for performance. The performance metrics include connection blocking, connection failure and path length. This research also proposes a novel heuristic link avoidance routing heuristics which is presented as an alternative to the others. Stochastic models resembling the realistic large scale traffic characteristics is used to quantitatively and qualitatively evaluate the survivability approaches under various load conditions. Independent single link failures are injected in order to examine the relative performance benefits in terms of survivability on imperfect networks.

In this research, Department of Energy's Energy Science Network (ESnet) is considered. ESnet contributes scientific researchers from more than 40 institutions and facilitates all DOE application networking needs. ESnet can support up to 100 Gbps nationwide and move datasets equal to 20 billion books every month. The network is managed by a centralized circuit reservation controller On-Demand Secure Circuits and Advance Reservation System (OSCARS) 1.0 prototype. OSCARS invokes appropriate Path Computation Engine (PCE) which computes paths based on traditional and tailored routing algorithms to schedule end to end circuit resources in advance of reservation establishment. This controller is extensively used in this research to analyze survivable algorithms for connection blocking, connection failures and path length.